---
title: "K means Clustering"
author: "ManojKumar Chalamala"
date: "10/29/2019"
output:
  html_document: default
  word_document: default
---

Install the required Packages
```{r}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("factoextra")
library(factoextra)
library(readr)
library(ISLR)
library(caret)
library(ggplot2)
```


Read the data

```{r}
Univ <- read.csv("Universities.csv")
str(Univ)
```

# a. Remove all records with missing measurements from the dataset.

```{r}
# remove all the missing values from the dataset by using na.omit
MyData <- na.omit(Univ)

# remove all the categorical variables from the dataset
MyData <- MyData[, c(-1, -2, -3)]
```

# b. For all the continuous measurements, run K-Means clustering. Make sure to normalize the measurements. How many clusters seem reasonable for describing these data? What was your optimal K?

```{r}
# Scaling the data frame (z-score) 
MyData <- scale(MyData)
View(MyData)
distance <- get_dist(MyData)
fviz_dist(distance)

# Determining K
# WSS method - Let us use an “elbow chart” to determine k
fviz_nbclust(MyData, kmeans, method = "wss")
# The chart shows that the elbow point 3 provides the best value for k. 

# Silhouette Method - Let us now apply the Silhouette Method to determine the number of clusters
fviz_nbclust(MyData, kmeans, method = "silhouette")
# We see that 3 is the ideal number of clusters. 

# Run the Kmeans algorithm for clustering. 
# We will choose an initial value of k = 3, number of restarts =15
k3 <- kmeans(MyData, centers = 3, nstart = 15)

# Visualize the output
k3$centers # output the centers

k3$size #No. of Universities in each cluster

k3 # mean of each cluster

k3$cluster[325] # Identify the cluster of the 325th observation as an example

fviz_cluster(k3, data = MyData) # Visualize the output

```

# c. Compare the summary statistics for each cluster and describe each cluster in this context (e.g., “Universities with high tuition, low acceptance rate...”)

```{r}
# mean of each cluster
k3$centers 
# Based on mean of each cluster compare the statistics of each cluster and described each cluster with High, low and medium context
ComparedData <- read.csv("ClusterComparison.csv") # Comparison of each cluster is stored in this excel
ComparedData #Displays the comparison of each cluster for all columns 

# plot an empty scatter plot
windows()
plot(c(0), xaxt = 'n', ylab = "", type = "l",
ylim = c(min(k3$centers), max(k3$centers)), xlim = c(0, 18))
# label x-axes
axis(1, at = c(1:17), labels = names(k3$centers))
# plot centroids
for (i in c(1:3))
lines(k3$centers[i,], lty = i, lwd = 2, col = ifelse(i %in% c(1, 2, 3),
"black", "dark grey"))
# name clusters
text(x = 0.5, y = k3$centers[, 1], labels = paste("Cluster", c(1:3)))

```

The above graph displays the comparison of summary stats of each cluster

# d. Use the categorical measurements that were not used in the analysis (State and Private/Public) to characterize the different clusters. Is there any relationship between the clusters and the categorical information?

```{r}
MyData1 <- Univ
MyData1 <- na.omit(MyData1)
View(MyData1)

MyData2 <- cbind(MyData1[,c(1:3)],k3$cluster)



MyData2$Public..1...Private..2. <- factor(MyData2$Public..1...Private..2.,levels = c(1,2),labels = c("public","private"))

MyData2$`k3$cluster` <- factor(MyData2$`k3$cluster`,levels = c(1,2,3),labels = c("Low ","Medium", "High"))

head(MyData2)

ggplot(MyData2, aes(x=MyData2$Public..1...Private..2., y=MyData2$State, color=MyData2$`k3$cluster`)) + geom_point()

```


Based on the graph,

Cluster 1 shows equaivalent proportion of both Public and Private Universities

Cluster 2 shows more Public Universities when compared with private

Cluster 3 shows more Private Universities when compared with public

# e. What other external information can explain the contents of some or all of these clusters?


```{r}
MyData2[MyData2$`k3$cluster`=="Low ",]
# Cluster 1 shows equaivalent proportion of both Public and Private Universities

MyData2[MyData2$`k3$cluster`=="Medium",]
# Cluster 2 shows more Public Universities when compared with private

MyData2[MyData2$`k3$cluster`=="High",]
# Cluster 3 shows more Private Universities when compared with public

k3$withinss

k3$iter

```

# f. Consider Tufts University, which is missing some information. Compute the Euclidean distance of this record from each of the clusters that you found above (using only the measurements that you have). Which cluster is it closest to? Impute the missing values for Tufts by taking the average of the cluster on those measurements. 

```{r}
MyData3 <- read.csv("Universities.csv")
# Considering Tufts University
MyData4 <- MyData3[ MyData3$College.Name== "Tufts University",]
View(MyData4)
# Compute Euclidian Distance
dist(rbind(MyData4[, -c(1, 2, 3,10)], k3$centers[1,]))
dist(rbind(MyData4[, -c(1, 2, 3,10)], k3$centers[2,]))
dist(rbind(MyData4[, -c(1, 2, 3,10)], k3$centers[3,]))
```

Based on the above distances it is closest to Cluster 3

Impute the missing values for Tufts by taking the average of the cluster on those measurements
```{r}
MyData5 <- read.csv("Universities.csv")
MyData5 <- na.omit(MyData5)
MyData5 <- cbind(MyData5,k3$cluster)
View(MyData5)
#Filter the values which falls under cluster3
MyData6 <- filter(MyData5, MyData5$`k3$cluster` == 3)
MyData7 <- mean(MyData6[,c(10)])
MyData7
MyData4$X..PT.undergrad <- MyData7
# After filling missing value with mean
MyData4
```

